---
title: "case study"
author: "Xiaoyu Wu"
date: "2023-10-10"
output: html_document
---


```{r, initial setup}
library(tidyverse)
library(ggridges)
library(leaflet)
knitr::opts_chunk$set(
  fig.width=6,
  fig.asp= .6,
  out.width="90%"
)

theme_set(theme_minimal()+theme(legend.position="bottom"))

options(
  ggplot2.cotinuous.color="viridis",
  ggplot2.cotinuous.fill="viridis"
)
# overwrite ggplot default coloring and take care of call the colors and fills 

scale_color_discrete=scale_color_viridis_d
scale_fill_discrete=scale_fill_viridis_d
```
Get the Data 
```{r}
library(p8105.datasets)
data(nyc_airbnb)

nyc_airbnb=
  nyc_airbnb |> 
  rename(borough=neighbourhood_group) |> 
  mutate(stars=review_scores_location/2)
```

Brainstorm questions
-Where are AirBNBs expensive?

```{r}
nyc_airbnb |>
  count(borough)
```

```{r}
price_rank=
  nyc_airbnb |>
  group_by(borough,room_type) |>
  summarize(median_price=median(price)) |>
  pivot_wider(
    names_from = room_type,
    values_from = median_price
  )
  
```

```{r}
nyc_airbnb |>
  filter(price<1000,room_type=="Entire home/apt") |>
  ggplot(aes(x=price))+
  geom_histogram()+
  facet_grid(.~borough)
```

```{r}
nyc_airbnb |>
  filter(price<1000,
         room_type=="Entire home/apt",
         borough=="Manhattan") |> 
  group_by(neighbourhood) |> 
  summarize(
    n_apt=n(),
    mean_price=mean(price)) |> 
  arrange(desc(mean_price))
```

```{r}
nyc_airbnb |>
  filter(price<1000,
         room_type=="Entire home/apt",
         borough=="Manhattan") |> 
  mutate(neighbourhood=fct_reorder(neighbourhood,price)) |> 
  ggplot(aes(x=price,y=neighbourhood)) +
  geom_density_ridges() 
```

## Try to make a map
```{r}
nyc_airbnb |>
  filter(price<1000,
         room_type=="Entire home/apt",
         borough=="Manhattan") |> 
  ggplot(aes(x=lat,y=long,color=price)) +
  geom_point()
```
Leaflet map:need html output to be interactive 
```{r}
nyc_airbnb |>
  filter(price<1000,
         room_type=="Entire home/apt",
         borough=="Manhattan") |>
  sample_n(500)|> 
  leaflet() |> 
  addProviderTiles(providers$CartoDB.Position) |> 
  addCircleMarkers(~lat,~long,radius=.5)
```

-How long are AirBNBs active?
-Are AirBNBs illegal and do they get shut down?
-Which units have the most availability?
-How is review score impacted by location?
-How many apts are run by one host?
   -does that affect price?
   -does that affect by neighbourhood?
-Where are AirBNBS?
   -Borough? Neighbourhood?
   -
