---
title: "reading_data_from_the_web"
author: "Xiaoyu Wu"
date: "2023-10-08"
output: github_document
---

```{r}
library(tidyverse)
library(rvest)
library(httr)
```
## Scrape a table 
I want the first table from this page

read in the html 
```{r}
url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"
drug_use_html = read_html(url)
# function from rvest
drug_use_html
```
extract the table(s)
```{r}
drug_use_html %>%
# html_nodes(css="table")
  html_table()
```
get the contents from the first list element
```{r}
table_marj = 
  drug_use_html %>% 
  html_table() %>% 
  first()
```
remove the first row
```{r}
table_marj = 
  drug_use_html %>% 
  html_table() %>% 
  first() %>%
  slice(-1) 

table_marj
```

learning assessment 

Create a data frame that contains the cost of living table for New York
```{r}
nyc_cost = 
  read_html("https://www.bestplaces.net/cost_of_living/city/new_york/new_york") %>%
  html_table(header = TRUE) %>%
  first()
```

## CSS Selectors Star Wars Movie Info 

I want the data from [here]("https://www.imdb.com/list/ls070150896/")

```{r}
url= "https://www.imdb.com/list/ls070150896/"
swm_html = read_html(url)
```
Grab elements that I want 
```{r}
title_vec = 
  swm_html %>%
  html_elements(".lister-item-header a") %>%
  html_text()

gross_rev_vec = 
  swm_html %>%
  html_elements(".text-small:nth-child(7) span:nth-child(5)") %>%
  html_text()

runtime_vec = 
  swm_html %>%
  html_elements(".runtime") %>%
  html_text()

swm_df = 
  tibble(
    title = title_vec,
    gross_rev = gross_rev_vec,
    runtime = runtime_vec)
```

## Get fomr Water Data 
This is from an API

Import this as a CSV and parse it
```{r}
nyc_water = 
  GET("https://data.cityofnewyork.us/resource/ia2d-e54m.csv") %>% 
  content("parsed")
# parse from table in html to tibble 

nyc_water=
  GET("https://data.cityofnewyork.us/resource/ia2d-e54m.json") %>% 
  content("text") %>% 
  jsonlite::fromJSON() %>% 
  as_tibble()
# can request API as a json file type 
```

```{r}

```

