---
title: "strings_and_factors"
author: "Xiaoyu Wu"
date: "2023-10-11"
output: github_document
---

```{r setup}
library(tidyverse)
library(rvest)
library(p8105.datasets)
knitr::opts_chunk$set(
  fig.width=6,
  fig.asp= .6,
  out.width="90%"
)

theme_set(theme_minimal()+theme(legend.position="bottom"))

options(
  ggplot2.cotinuous.color="viridis",
  ggplot2.cotinuous.fill="viridis"
)
# overwrite ggplot default coloring and take care of call the colors and fills 

scale_color_discrete=scale_color_viridis_d
scale_fill_discrete=scale_fill_viridis_d
```

## Strings and manipulation
use str_detect to find cases where the match exists 
```{r}
string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "jeff")
```

use str_replace to replace an instance of a match with something else
```{r}
str_replace(string_vec, "jeff", "Jeff")
```

designate matches at the beginning of a line
```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "^i think")
# beginning of the line 
```

designate matches at the end of a line
```{r}
str_detect(string_vec, "i think$")
# end of the line 
```

designate a list of characters that will count as a match
```{r}
string_vec = c(
  "Y'all remember Pres. HW Bush?",
  "I saw a green bush",
  "BBQ and Bushwalking at Molonglo Gorge",
  "BUSH -- LIVE IN CONCERT!!"
  )

str_detect(string_vec,"[Bb]ush")
```

provide a range of letters or numbers that count as a match
```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "^[0-9][a-zA-Z]")
# here capital or lower case letter are both ok followed by a number
```

The character "." matches anything
```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")
# actual literal dot we use "\\."; we put "\\" to specify the true dot here 
```

Some characters are “special”. These include [ and ], ( and ), and .
```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[")
# indicate they’re special using \
```
## Factors

```{r}
vec_sex = factor(c("male", "male", "female", "female"))
vec_sex
```

```{r}
as.numeric(vec_sex)
```

```{r}
vec_sex = fct_relevel(vec_sex, "male")
# make male the first category or level that corresponds to 1
vec_sex
```

```{r}
as.numeric(vec_sex)
```

## NSDUH
load the data
```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

table_marj = 
  read_html(nsduh_url) %>% 
  html_table() %>% 
  first() %>%
  slice(-1)
```
tidy the data
```{r}
data_marj = 
  table_marj %>%
  select(-contains("P Value")) %>%
# get rid of every column that contains the string "p value" 
  pivot_longer(
    -State,
# everything that isn't in the state column 
    names_to = "age_year", 
    values_to = "percent") %>%
  separate(age_year, into = c("age", "year"), sep = "\\(") %>%
#  we split age and year at the open parentheses using "\\(" and delete the seperate element"(". 
  mutate(
    year = str_replace(year, "\\)", ""),
# we stripped out the close parenthesis in mutate
    percent = str_replace(percent, "[a-c]$", ""),
# to remove character superscripts, we replaced any character using "[a-c]$"
    percent = as.numeric(percent)) %>%
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
# don't want my state variable to be one of "Total U.S.", "Northeast", "Midwest", "South", "West". 
```
make a plot
```{r}
data_marj %>%
  filter(age == "12-17") %>% 
  mutate(State = fct_reorder(State, percent)) %>% 
  ggplot(aes(x = State, y = percent, color = year)) + 
    geom_point() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## NYC Restaurant Inspections

```{r}
data("rest_inspec")

rest_inspec %>% 
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  pivot_wider(names_from = grade, values_from = n)
```

```{r}
rest_inspec =
  rest_inspec %>%
  filter(grade %in% c("A", "B", "C"), boro != "Missing") %>% 
# remove inspections with scores other than A, B, or C, and also remove the restaurants with missing boro information
  mutate(boro = str_to_title(boro))
# clean up boro names
```
Let’s focus only on pizza places for now, and re-examine grades by borough

Not right due to case sensitive 
```{r}
rest_inspec %>% 
  filter(str_detect(dba, "Pizza")) %>% 
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  pivot_wider(names_from = grade, values_from = n)
```
Right Version 
```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>% 
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  pivot_wider(names_from = grade, values_from = n)
```

Visualization 
```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>%
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar() 
```
have things in a different order – maybe number of pizza places
```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>%
  mutate(boro = fct_infreq(boro)) %>%
# reorder factor levels by number of observations with each level (largest first)
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar() 
```
A dedicated function for renaming factor levels
```{r}
rest_inspec %>% 
  filter(str_detect(dba, regex("pizza", ignore_case = TRUE))) %>%
  mutate(
    boro = fct_infreq(boro),
    boro = fct_recode(boro, "The City" = "Manhattan")) %>%
  ggplot(aes(x= boro, fill = grade)) +
  geom_bar()
# using fct_recode 
```

## Weather Data 

Load the data 
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

```{r}
weather_df
```

```{r}
weather_df %>%
  mutate(name = forcats::fct_relevel(name, c("Waikiki_HA", "CentralPark_NY", "Waterhole_WA"))) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```

```{r}
weather_df %>%
  mutate(name = forcats::fct_reorder(name, tmax)) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```

## Linear regression 
```{r}
weather_df %>%
  lm(tmax ~ name, data = .)
# fit tmax as our outcome and name as our predictor 
```

```{r}
weather_df %>%
  mutate(name = forcats::fct_relevel(name, c("Waikiki_HA", "CentralPark_NY", "Waterhole_WA"))) %>% 
  lm(tmax ~ name, data = .)
```

